---
title: "Homework 2 solutions"
author: "Yunxi Yang"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

### Problem 1

Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables.

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

### Problem 2

```{r First read and clean the Mr. Trash Wheel collection data sheet}
mr_trash_wheel_df = 
  read_excel("Trash Wheel Collection Data.xlsx", range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = as.integer(sports_balls),
         year = as.integer(year)) 
mr_trash_wheel_df
```



```{r read and clean the Professor Trash Wheel for P2}
pro_trash_df = 
  read_excel("./Data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Professor Trash Wheel", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = as.integer(round(sports_balls, 0)),
         year = as.character(year),
         name = "professor_trash_wheel")
```

```{r Then read and clean the Professor Trash Wheel collection data sheet}
prof_trash_wheel_df = 
  read_excel("Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = cell_cols("A:M")) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = as.integer(year)) 
prof_trash_wheel_df
```

```{r combine these two dataset}
combined_trash_df = bind_rows(mr_trash_wheel_df, prof_trash_wheel_df)
combined_trash_df
```

Descriptions about these data: 
1. There are `r colnames(combined_trash_df)` variables. 
2. There are totally `r nrow(combined_trash_df)` rows of observations, and `r ncol(combined_trash_df)` columns of observations.
3. The total weight of trash collected by Professor Trash Wheel is `r sum(prof_trash_wheel_df$weight_tons)` tons. 
4. The total number of sport balls collected by Mr. Trash Wheel in 2020 is `r sum(mr_trash_wheel_df$sports_balls, mr_trash_wheel_df$year == 2020)` balls. 


### Problem 3

```{r import datasets of pols_month, unemployment and snp}
pols_month_df = read_csv("fivethirtyeight_datasets/pols-month.csv")
unemployment_df = read_csv("fivethirtyeight_datasets/unemployment.csv")
snp_df = read_csv("fivethirtyeight_datasets/snp.csv")
```
